Design and Implementation of a Multithreaded Virtual Machine for Executing Linear Logic Programs

### 2014-08-12 paper.tex
update

Virtual machines are a popular technique for implementing interpreters
for high level programming languages.  Due to the increased
availability of parallel machines and distributed architectures,
several machine models have been developed with parallelism in
mind~\cite{Kara:1997:AMM:265274}.  One example of such machine is the
Parallel Virtual Machine (PVM)~\cite{Sunderam90pvm:a}, which serves as
an abstraction to program heterogeneous computers as a single
machine. Another important machine is the Threaded Abstract Machine
(TAM)~\cite{CullerGSvE93,goldstein-tr94}, which defines a
self-scheduled machine language of parallel threads where a program is
represented as conventional control flow.

Prolog, the most prominent logic programming language, has a rich
history of virtual machine research centered around the Warren
Abstract Machine (WAM)~\cite{AICPub641:1983}. 
Prolog is naturally parallel because several clauses for the same goal
(AND-parallelism) or all goals in a clause (OR-parallelism) can be
tried in parallel. Different abstract machines for AND-parallelism has
been developed on top of the
WAM~\cite{Hermenegildo:1986:AMB:913061,Lin:1988:AEL:900478}.  For
OR-parallelism we have several models such as: the SRI
model~\cite{Warren:1987:OEM:67683.67699}, the Argonne
model~\cite{ButlerDLOOS88}, the MUSE model~\cite{Ali:1990fk} and the
BC machine~\cite{Ali88}. While all those models have been developed
using the WAM, there are also some parallel machines totally different
from the WAM, such as the PPAM~\cite{Kacsuk:1990:EMP:533578}, which is
based on a data-flow model.

Several linear logic programming languages have been developed in the past~\cite{Miller85anoverview}.
Lolli, a programming language based on a fragment of intuitionistic linear logic~\cite{Hodas94logicprogramming}, 
proves goals by lazily managing the context of linear resources during backward-chaining proof search.
Forum~\cite{Miller94} is an extension of Lolli that is based on a classical proof system. While Lolli uses an intuitionistic proof system that supports only one conclusion, Forum allows multiple conclusions. Forum is therefore better adapted to represent concurrent computations than Lolli, since many multiple goals need to be proved at once.

Another extension of Lolli is LolliMon~\cite{Lopez:2005:MCL:1069774.1069778}, a concurrent linear logic programming language that integrates
both forward and backward-chaining search, where the backward-chaining phase is done sequentially but
the forward-chaining is done concurrently inside a monad. The backward-chaining phase is suspended
between the forward-chaining phases, where a fix-point is computed.
LolliMon is derived from the concurrent logical framework called CLF~\cite{Watkins:2004uq}.

@@

@@

\bibitem[Butler et~al.(1988)Butler, Disz, Lusk, Olson, Overbeek, and
  Stevens]{ButlerDLOOS88}
R.~Butler, T.~Disz, E.~L. Lusk, R.~Olson, R.~A. Overbeek, and R.~L. Stevens.
\newblock Scheduling or-parallelism: An argonne perspective.
\newblock In R.~A. Kowalski and K.~A. Bowen, editors, \emph{ICLP/SLP}, pages
  1590--1605. MIT Press, 1988.
\newblock ISBN 0-262-61056-6.
\newblock URL
  \url{http://dblp.uni-trier.de/db/conf/iclp/iclp88.html#ButlerDLOOS88}.

>>>>

Virtual machines are a popular technique for implementing interpreters
for high level programming languages.  Due to the increased
availability of parallel machines and distributed architectures,
several machine models have been developed with parallelism in
mind~\cite{Kara:1997:AMM:265274}.  One example is the
Parallel Virtual Machine (PVM)~\cite{Sunderam90pvm:a}, which serves as
an abstraction to program heterogeneous computers as a single
machine. Another important machine is the Threaded Abstract Machine
(TAM)~\cite{CullerGSvE93,goldstein-tr94}, which defines a
self-scheduled machine language of parallel threads where a program is
represented as conventional control flow.

Prolog, the most prominent logic programming language, has a rich
history of virtual machine research centered around the Warren
Abstract Machine (WAM)~\cite{AICPub641:1983}. 
Prolog is naturally parallel because several clauses for the same goal
(AND-parallelism) or all goals in a clause (OR-parallelism) can be
tried in parallel. Different abstract machines for AND-parallelism has
been developed for the
WAM~\cite{Hermenegildo:1986:AMB:913061,Lin:1988:AEL:900478}.  For
OR-parallelism we have several models such as: the SRI
model~\cite{Warren:1987:OEM:67683.67699}, the MUSE model~\cite{Ali:1990fk} and the
BC machine~\cite{Ali88}. Although these models are built on top of the WAM,
some machines use different approaches, such as the
PPAM~\cite{Kacsuk:1990:EMP:533578} with its
data-flow model.

Several linear logic programming languages have been developed in the past~\cite{Miller85anoverview}.
Lolli, a programming language based on a fragment of intuitionistic linear logic~\cite{Hodas94logicprogramming}, 
proves goals by lazily managing the context of linear resources during backward-chaining proof search.
Forum~\cite{Miller94} is an extension of Lolli that is based on a classical proof system. While Lolli uses an intuitionistic proof system that supports only one conclusion, Forum allows multiple conclusions. Forum is therefore better adapted to represent concurrent computations than Lolli, since many multiple goals need to be proved at once.
Another extension of Lolli is LolliMon~\cite{Lopez:2005:MCL:1069774.1069778}, a concurrent linear logic programming language that integrates
both forward and backward-chaining search, where the backward-chaining phase is done sequentially but
the forward-chaining is done concurrently inside a monad. The backward-chaining phase is suspended
between the forward-chaining phases, where a fix-point is computed.
LolliMon is derived from the logical framework called CLF~\cite{Watkins:2004uq}.

@@

We thank the anonymous reviewers for their insightful comments and suggestions
that helped us improve the quality of the paper.

@@
